Biologically-Constrained 3D Agglomeration for Connectome Reconstruction

### 2018-03-14 papers/eccv2018/0_introduction.tex
donglai's pass on abstract-intro-related

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{./figures/teaser_v2.png}
	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation.}
	\label{fig:teaser}
\end{figure}

The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{haehn2017scalable,kasthuri2015saturated}. 
Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) allow researchers to produce terabytes of image data per hour~\cite{hildebrand2017whole}. 
It is infeasible for domain experts to manually reconstruct this vast amount of image data~\cite{haehn2014design}. 
State-of-the-art automatic reconstruction approaches use pixel-based segmentation with convolutional neural networks (CNNs) followed by agglomeration strategies~\cite{seymour2016rhoananet,lee2015recursive,nunez2014graph,parag2017anisotropic,ronneberger2015u,zlateski2015image}.
These \textit{bottom-up pixel-based} methods produce excellent results but still fall short of acceptable error rates for large volumes.

We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
Our method imposes both local and global biological constraints onto the output segmentation to match the underlying structure of neuronal processes.
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser}a). 
Next, we train a CNN classifier to learn biological constraints on the local structural shape of a neuron (Fig.~\ref{fig:teaser}b). 
This classifier produces probabilities that two segments belong to the same neuron based on the input shapes of the segments.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific topological constraints from biology (Fig.~\ref{fig:teaser}c).

Our approach operates at a level of abstraction above existing pixel-based methods. 
This allows us to leverage both local and global information to produce more accurate reconstructions. 
Our method is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining.
Images generated by electron microscopes often differ in appearance because of variations in staining techniques \cite{briggman2012volume}.
By removing the dependence of our algorithm on the input images, we greatly reduce the need for additional costly ground truth data for each new stack of images.
Using the graph induced by the segmentation allows us to enforce global biological constraints on the reconstruction. 
Our dual approach of assessing local decisions in a global context yields accuracy improvements over existing reconstruction methods.

This work makes the following contributions: (1) a novel top-down framework that extracts graphs from skeletonized 3D segmentations for improved neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors; (3) an empirical evaluation of our method on several connectomics datasets; (4) improved performance over a state-of-the-art pixel-based reconstruction approach on average by \TODO{XX}\% with suitable running times.

>>>>

% \begin{figure}[t!]
% 	\centering
% 	\includegraphics[width=\linewidth]{./figures/teaser_v2.png}
% 	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation.}
% 	\label{fig:teaser}
% \end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{./figures/bc/teaser_bc_v2.pdf}
	\caption{Biological constraints for 3D segments in Connectomics.}
	\label{fig:bc}
\end{figure}

The field of connectomics is concerned about reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{haehn2017scalable,kasthuri2015saturated}. 
%Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) allow researchers to produce terabytes of image data per hour~\cite{hildebrand2017whole}. 
%It is infeasible for domain experts to manually reconstruct this vast amount of image data~\cite{haehn2014design}. 
State-of-the-art automatic 3D reconstruction approaches use voxel-based convolutional neural networks (CNNs) and watershed transform to generate initial segmentation, followed by agglomeration steps~\cite{seymour2016rhoananet,lee2015recursive,nunez2014graph,parag2017anisotropic,ronneberger2015u,zlateski2015image}.
%These \textit{bottom-up pixel-based} methods produce excellent results but still fall short of acceptable error rates for large volumes.
For the agglomeration, most methods do not make use of the biological knowledge of the 3D shape of neural structures.
Reported on small-sized volumes, a simple threshold-based mean agglomeration method can achieve superhuman result~\cite{lee2017superhuman}.
However, it is unclear if it is the case for larger volumes.

%\TODO{the selling point is to add biological constraint}
%We present a \textit{top-down graph-based} method that builds on the outputs of bottom-up pixel-based segmentation approaches. 
We here present a new agglomeration method that imposes both local and global biological constraints onto the output segmentation to match the underlying structure of neuronal processes.
We first extract skeletons from labels in the input segmentation and generate a simplified graph (Fig.~\ref{fig:teaser}a). 
Next, we train a CNN classifier to learn biological constraints on the local structural shape of a neuron (Fig.~\ref{fig:teaser}b). 
This classifier produces probabilities that two segments belong to the same neuron based on the input shapes of the segments.
These probabilities correspond to edge weights in our graph extracted from the input segmentation.
We then use a graph optimization algorithm to partition the graph into a final improved reconstruction by enforcing domain-specific topological constraints from biology (Fig.~\ref{fig:teaser}c).

%Our approach operates at a level of abstraction above existing pixel-based methods. 
%This allows us to leverage both local and global information to produce more accurate reconstructions. 
With the biological constraints, we can make use of both geometric and topological prior information to produce more accurate reconstructions.
\TODO{refer to figure 1}
Our method is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining.
Images generated by electron microscopes often differ in appearance because of variations in staining techniques \cite{briggman2012volume}.
By removing the dependence of our algorithm on the input images, we greatly reduce the need for additional costly ground truth data for each new stack of images.
Using the graph induced by the segmentation allows us to enforce global biological constraints on the reconstruction. 
Our dual approach of assessing local decisions in a global context yields accuracy improvements over existing reconstruction methods.

This work makes three contributions.
First, a novel top-down framework that extracts graphs from skeletonized 3D segmentations for improved neural reconstruction of connectomics data; 
Then, a region-based CNN classifier to detect errors; 
%(3) an empirical evaluation of our method on several connectomics datasets;  % this is not contribution
Last, improved performance over a state-of-the-art pixel-based reconstruction approach on average by \TODO{XX}\% with suitable running times.

### 2018-03-14 papers/eccv2018/1_related_work.tex
donglai's pass on abstract-intro-related

\paragraph{Pixel-based methods.}
A significant amount of connectomics research considers the problem of extracting segmentation information at the pixel (i.e., voxel) level from the raw EM images. 
Some methods use a random forest populated with hand-designed features based on image characteristics around a given pixel to predict membrane probabilities~\cite{kaynig2015large}.
Most recent advancements use convolutional neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
Originally, networks predicted membrane probabilities per each image slice using 2D convolutions~\cite{seymour2016rhoananet,ronneberger2015u,ciresan2012deep,jain2010boundary,kaynig2015large,amelio_segmentation}.
However, recent research extends these convolutions to include the z-dimension improving performance~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional}.
Many of these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). 
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. 
These networks produce impressive accuracies but at a high computational cost.
Data augmentation in both training and testing can increase the prediction accuracy in boundary detection~\cite{zeng2017deepem3d}.
Lee et al. surpass the estimated human accuracy on the SNEMI3D dataset with a variant of the 3D U-Net architecture~\cite{lee2017superhuman}.

\paragraph{Region-based methods.}

Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}. 
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.
Many methods build on top of these region-based strategies and train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.

\paragraph{Error-correction methods.}
Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

>>>>

% \begin{figure}[t]
% 	\subfloat[]{
% 		\includegraphics[width=.47\linewidth]{figures/schema/pre-multicut-with-arrows.pdf}
% 	}
% 	\subfloat[]{
% 		\includegraphics[width=.47\linewidth]{figures/schema/post-multicut.png}
% 	}
% 	\caption{Example improvement of neural reconstruction. (a) We extract 3D skeletons from pixel-based segmentation algorithms to create a graph representation. Red arrows highlight errors in the input segmentation. (b) We improve the segmentation accuracy using a graph partitioning algorithm, leveraging both local and global information.}
% 	\label{fig:improved-reconstruction}
% \end{figure}

\subsubsection{Initial Segmentation}
A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level from raw EM images.

First, intermediate representations like boundary or affinity maps are generated. Some methods use a random forest populated with hand-designed features based on image characteristics around a given pixel to predict membrane probabilities~\cite{kaynig2015large}.
Most recent advancements use convolutional neural networks since machine-learned features outperform hand-designed ones~\cite{bogovic2013learned}.
Originally, networks predicted membrane probabilities per each image slice using 2D convolutions~\cite{seymour2016rhoananet,ronneberger2015u,ciresan2012deep,jain2010boundary,kaynig2015large,amelio_segmentation}.
However, recent research extends these convolutions to include the z-dimension improving performance~\cite{lee2015recursive,parag2017anisotropic,cciccek20163d,turaga2010convolutional}.
Many of these networks produce probabilities for the affinity between two voxels (i.e., the probability that adjacent voxels belong to the same neuron). 
The MALIS cost function is specifically designed for generating affinities that produce good segmentations~\cite{briggman2009maximin}. 
More recently, flood-filling networks produce segmentations by training an end-to-end neural network that goes from EM images directly to label volumes~\cite{januszewski2016flood}. 
These networks produce impressive accuracies but at a high computational cost.
Data augmentation in both training and testing can increase the prediction accuracy in boundary detection~\cite{zeng2017deepem3d}.
Lee et al. surpass the estimated human accuracy on the SNEMI3D dataset with a variant of the 3D U-Net architecture~\cite{lee2017superhuman}.

Then, these intermediate representations are transformed into segmentation.
Some early techniques apply computationally expensive graph partitioning algorithms with a single node per superpixel~\cite{andres2012globally}.
Several pixel-based approaches generate probabilities that neighboring pixels belong to the same neuron.
Often a watershed algorithm will then cluster pixels into super-pixels~\cite{zlateski2015image}.

\subsubsection{Agglomeration}
Given the initial segmentation, some methods train random-forest classifiers to produce the final segmentations~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,10.1371/journal.pone.0125825}.
Some methods rely solely on the accuracy of the predicted affinities and agglomerate hierarchically based on the mean affinity between supervoxels~\cite{lee2017superhuman,funke2017deep}.
Pape et al. present a scalable multicut algorithm to partition superpixels~\cite{beier2017multicut}.
\TODO{add learning-based algorithm} ~\cite{jain2011learning,bogovic2013learned}
%\subsubsection{Error-correction methods}
%Additional research builds on top of these region-based methods to correct errors in the segmentation either using human proofreading~\cite{haehn2014design,haehn2017guided,mojo2} or automatic methods~\cite{rolnick2017morphological,error_correction_using_CNN}.
%However, to our knowledge, our method is the first to extract a 3D graph from an input segmentation for a true top-down\TODO{?} error correction approach. 
%This allows us to enforce domain-specific biology constraints and use efficient graph partitioning algorithms. 
% Many segmentation and clustering algorithms use graph partitioning techniques~\cite{andres2012globally} or normalized cuts for traditional image segmentation~\cite{kappes2016higher,shi2000normalized,tatiraju2008image}.
% Even though graph partitioning is an NP-Hard problem~\cite{demaine2006correlation} there are several useful multicut heuristics that provide good approximations with reasonable computational costs~\cite{horvnakova2017analysis}. 
%We use the method of Keuper et al.~\cite{keuper2015efficient} to partition the extracted graph into the final neural reconstruction.

### 2018-03-14 papers/eccv2018/2_method.tex
donglai's pass on abstract-intro-related

\begin{figure*}[t!]
	\begin{center}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
	\end{center}
	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
	\label{fig:overview}
\end{figure*}

There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
In a split error, two segments should have been merged into one neuronal process. 
In a merge error, one segment corresponds to more than one neuron.
Generally, it is much more difficult to correct merge errors than to correct split errors,
as the space of possible split proposals grows quickly~\cite{parag2015properties}.
Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
Our method takes as input over-segmentations of EM image volumes generated by scalable, state-of-the-art connectomics reconstruction pipelines (Sec.~\ref{sec:neuroproof}). 
Our goal is to identify locations of split errors and merge the corresponding segments automatically.

@@

\begin{figure}[t]
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Two example candidates for merging. The two segments in the left example do not belong to the same neuron as evidenced by the sharp turning radius.}
	\label{fig:turn-radii}
\end{figure}

@@

\subsection{Edge Weights Assignment}
\label{sec:edge-weights}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
We train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.

@@

\begin{figure}[t]
	\centering
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error2.png}				
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error3.png}
	\end{minipage}
	\caption{Three erroneously split segments.}
	\label{fig:merge_candidates}
\end{figure}

@@

\subsubsection{Lifted Multicut}

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{./figures/teaser_v2.png}
	\caption{Our proposed framework uses both local and global biological constraints to improve an input segmentation.}
	\label{fig:teaser}
\end{figure}
% \begin{figure*}[t!]
% 	\begin{center}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
% 	\end{center}
% 	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
% 	\label{fig:overview}
% \end{figure*}

%There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
%In a split error, two segments should have been merged into one neuronal process. 
%In a merge error, one segment corresponds to more than one neuron.
%Generally, it is much more difficult to correct merge errors than to correct split errors, as the space of possible split proposals grows quickly~\cite{parag2015properties}.
%Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
Our method takes as input over-segmentations of EM image volumes generated by scalable, state-of-the-art connectomics reconstruction pipelines (Sec.~\ref{sec:neuroproof}). 
Our goal is to identify locations of split errors and merge the corresponding segments automatically.

@@

@@

\begin{figure}[t]
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Two example candidates for merging. The two segments in the left example do not belong to the same neuron as evidenced by the sharp turning radius.}
	\label{fig:turn-radii}
\end{figure}
\subsection{Edge Weights Assignment}
\label{sec:edge-weights}
We assign edge weights $w_e$ to each edge where the weight corresponds to the probability that two nodes belong to the same neuron.
We train a 3D CNN classifier to learn from the manually labeled oversegmentation input volume (Sec.~\ref{sec:dataset}).
If the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}$, where $\beta$ is a tunable parameter that encourages over- or under-segmentation.

@@

@@

\begin{figure}[t]
	\centering
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error2.png}				
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error3.png}
	\end{minipage}
	\caption{Three erroneously split segments.}
	\label{fig:merge_candidates}
\end{figure}
\subsubsection{Lifted Multicut}

### 2018-03-14 papers/eccv2018/paper.tex
donglai's pass on abstract-intro-related

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
\title{Scalable Graph-Based Neural \\ Reconstruction for Connectomics}

@@

\begin{abstract}
Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
Current state-of-the-art automatic methods classify neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions.
However, these approaches widely neglect global geometric properties and other biological constraints that are inherent in the graph structure of neural wiring diagrams.
In this work, we follow pixel-based reconstruction by a biologically-constrained graph-based method to more accurately approximate neural pathways.
We first generate skeletons in 3D from the neuron labels of the pixel-based segmentation.
Next, we simplify this skeletonized 3D network into a graph with nodes corresponding to labels from the segmentation and edges identifying potential locations of split errors.
A CNN classifier produces edge weights for the graph corresponding to the probability of error. 
We formulate the graph into a lifted multicut partitioning problem to improve the final segmentation.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, skeletonization, deep learning, multicut}
\end{abstract}

>>>>

\authorrunning{ECCV-18 submission ID \ECCV18SubNumber}
%\title{Scalable Graph-Based Neural \\ Reconstruction for Connectomics}
\title{Biologically-Constrained 3D Agglomeration for Connectome Reconstruction}

@@

\begin{abstract}
% maybe better to cut to the point
%Advancements in electron microscopy image acquisition have created connectomics datasets terabytes in size making manual reconstruction of neuronal structures infeasible.
%In the past decade in Connectomics, much progress has been made to classify neural membranes or affinities from the raw pixel and to watershed transform them into initial segments.
%Current state-of-the-art automatic methods classify neural membranes or affinities from the raw pixel followed by agglomeration methods to create full neuron reconstructions.
%An agglomerative segmentation algorithm merges initial pair of clusters.
Current state-of-the-art connectome reconstruction pipelines have three major steps: affinity prediction from raw voxels, initial segmentation with watershed transform and refined segmentation with agglomeration.
%In the past deTraditionally, neuron segmentation from 3D electron microscopy (EM)
For the agglomeration, most methods neglect geometric and topological constraints that are inherent in the graph structure of neural wiring diagrams.

In this work, we use a biologically-constrained graph-based method to better reconstruct neural pathways.
Given the initial segmentation, we first create a graph with nodes corresponding to segment labels and edges identifying potential locations of split errors based on the skeleton structure of the segmentation.
Next, we train a convolutional neural network to capture the geometric properties of segment connections and produce edge weights as the probability of merging two nodes.
Finally, we formulate the agglomeration as a lifted multicut partitioning problem with topological constraints to avoid cycles.
We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average split variation of information improvement of \TODO{XX}\%.
\keywords{connectomics, biological constraints, agglomeration, skeletonization, deep learning, lifted multicut}
\end{abstract}
